Kids PM Summary

### 2019-12-05 data_cleaning_initial.Rmd
PM data cleaning updates

---
title: "KidsAIR initial data work"
author: "Ethan Walker"
date: "Started 19 Nov 2019, Updated 4 Dec 2019"
output: pdf_document
---

@@

# Load, format, and save KidsAIR Dusttrack data
```{r}
##### Read in dusttrack .txt files (from WMT location) #####

@@

# Format combined file from above; combine with other WMT/PM data below
wmt_pm_txt <- initial_data %>% 
  mutate(date = mdy(`MM/dd/yyyy`),
         time = hms(`hh:mm:ss`),
         pm = as.numeric(`mg/m^3`),
         area = "WMT") %>% 
  select(home_winter_id, date, time, pm, area)

##### Next, load other files from all locations #####

##### Load combined dusttrack files from Emily W #####
##### Compare these with the raw data pulled in above #####

>>>>

---
title: "KidsAIR initial data work"
author: "Ethan Walker"
date: "Started 19 Nov 2019, Updated 5 Dec 2019"
output: pdf_document
---

@@

# Load, format, and save KidsAIR Dusttrack data
```{r}
##### Read in dusttrack .txt files (from WMT) #####

@@

# Format combined file from above; combine with other PM data below
wmt_pm_txt <- initial_data %>% 
  mutate(date = mdy(`MM/dd/yyyy`),
         time = `hh:mm:ss`,
         pm = as.numeric(`mg/m^3`),
         area = "WMT") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  select(home_winter_id, datetime, date, time, pm, area)

##### Read in dusttrack .csv files (all areas) #####

### WMT ###
# List files
list_csv_files <- list.files("Input/WMT/dusttrack_csv_files")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/WMT/dusttrack_csv_files")
initial_data = tibble(file_list = list_csv_files) %>%
  extract(file_list, "home_winter_id", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
wmt_pm_csv <- initial_data %>% 
  mutate(date = mdy(Date),
         time = Time,
         pm = as.numeric(AEROSOL),
         area = "WMT") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  select(home_winter_id, datetime, date, time, pm, area) %>% 
  filter(!is.na(pm))

### NN ###
# List files
list_csv_files <- list.files("Input/NN/dusttrack_csv_files")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/NN/dusttrack_csv_files")
initial_data = tibble(file_list = list_csv_files) %>%
  extract(file_list, "home_winter_id", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
nn_pm_csv <- initial_data %>% 
  mutate(date = mdy(Date),
         time = Time,
         pm = as.numeric(AEROSOL),
         area = "NN") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  select(home_winter_id, datetime, date, time, pm, area) %>% 
  filter(!is.na(pm))

### AK ###
# List files
list_csv_files <- list.files("Input/AK/dusttrack_csv_files")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/AK/dusttrack_csv_files")
initial_data = tibble(file_list = list_csv_files) %>%
  extract(file_list, "home_winter_id", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
ak_pm_csv <- initial_data %>% 
  mutate(date = mdy(Date),
         time = Time,
         pm = as.numeric(AEROSOL),
         area = "AK") %>% 
  unite("datetime", c("date", "time"), sep = " ", remove = FALSE) %>% 
  select(home_winter_id, datetime, date, time, pm, area) %>% 
  filter(!is.na(pm))

##### Read in other dusttrack .csv files (WMT and NN) #####

### WMT ### 
# List files
list_other_files <- list.files("Input/WMT/dusttrack_other_files")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/WMT/dusttrack_other_files")
initial_data = tibble(file_list = list_other_files) %>%
  extract(file_list, "home_winter_id", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
wmt_pm_other <- initial_data %>% 
  mutate(trash1 = `Instrument Name`,
         trash2 = `DustTrak II`) %>% 
  select(home_winter_id, trash1, trash2) %>% 
  group_by(home_winter_id) %>% 
  mutate(trash3 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
                              trash2, "NA"),
         trash4 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
                              trash1, "NA"),
         trash1 = as.numeric(trash1),
         trash1 = as.character(trash1),
         trash5 = if_else(trash3 != "NA", trash3, trash1)) %>% 
  filter(!is.na(trash5)) %>% 
  replace_with_na(replace = list(trash3 = "NA", trash4 = "NA")) %>% 
  spread(trash4, trash3) %>% 
  mutate(`Test Start Time` = lag(`Test Start Time`)) %>% 
  unite("datetime", `Test Start Date`:`Test Start Time`, sep = " ") %>%
  mutate(datetime = lead(datetime)) %>% 
  replace_with_na(replace = list(datetime = "NA NA")) %>% 
  mutate(datetime2 = first(datetime),
         datetime = mdy_hms(datetime2),
         trash1 = as.numeric(trash1)) %>% 
  filter(!is.na(trash1)) %>% 
  mutate(datetime2 = datetime + seconds(trash1),
         datetime2 = ymd_hms(datetime2)) %>% 
  separate(datetime2, into = c("date", "time"), sep = " ", remove = FALSE) %>% 
  mutate(pm = as.numeric(trash2),
         datetime = datetime2,
         area = "WMT") %>% 
  select(home_winter_id, datetime, date, time, pm, area) %>% 
  ungroup()

### NN ### 
# List files
list_other_files <- list.files("Input/NN/dusttrack_other_files")

# Set working directory and load files in list; extract file name and add as column
## run next 6 lines together
setwd("Input/NN/dusttrack_other_files")
initial_data = tibble(file_list = list_other_files) %>%
  extract(file_list, "home_winter_id", remove = FALSE) %>%
  mutate(save_data = lapply(file_list, read_csv)) %>%
  unnest(save_data) %>%
  select(-file_list)

# Format combined file from above; combine with other PM data below
nn_pm_other <- initial_data %>% 
  mutate(trash1 = `Instrument Name`,
         trash2 = `DustTrak II`) %>% 
  select(home_winter_id, trash1, trash2) %>% 
  filter(!is.na(trash1)) %>% 
  group_by(home_winter_id) %>% 
  mutate(trash3 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
                              trash1, "NA"),
         trash4 = if_else(trash1 == "Test Start Date" | trash1 == "Test Start Time", 
                              trash2, "NA"),
         trash1 = as.numeric(trash1),
         trash1 = as.character(trash1),
         trash5 = if_else(trash3 != "NA", trash3, trash1)) %>% 
  replace_with_na(replace = list(trash3 = "NA", trash4 = "NA")) %>% 
  filter(!is.na(trash5)) %>% 
  spread(trash3, trash4) %>% 
  mutate(`Test Start Time` = lag(`Test Start Time`)) %>% 
  unite("datetime", `Test Start Date`:`Test Start Time`, sep = " ") %>%
  mutate(datetime = lead(datetime)) %>% 
  replace_with_na(replace = list(datetime = "NA NA")) %>% 
  mutate(datetime2 = first(datetime),
         datetime = mdy_hms(datetime2),
         trash1 = as.numeric(trash1)) %>% 
  filter(!is.na(trash1)) %>% 
  mutate(datetime2 = datetime + seconds(trash1),
         datetime2 = ymd_hms(datetime2)) %>% 
  separate(datetime2, into = c("date", "time"), sep = " ", remove = FALSE) %>% 
  mutate(pm = as.numeric(trash2),
         datetime = datetime2,
         area = "NN") %>% 
  select(home_winter_id, datetime, date, time, pm, area) %>% 
  ungroup() %>% 
  mutate(home_winter_id = if_else(home_winter_id == "220001" | home_winter_id == "220002",
                                  "22", home_winter_id),
         home_winter_id = if_else(home_winter_id == "550001" | home_winter_id == "550002",
                                  "55", home_winter_id),
         home_winter_id = if_else(home_winter_id == "810001" | home_winter_id == "810002",
                                  "81", home_winter_id))

##### Bind all files from above #####
kids_pm <- rbind(wmt_pm_csv, wmt_pm_txt, wmt_pm_other,
                nn_pm_csv, nn_pm_other,
                ak_pm_csv)

# Save pm dataset as RDS and CSV
write_rds(kids_pm, "Output/kids_pm_full.rds")
#write_csv(kids_pm, "Output/kids_pm_full.csv")

# Load data and clean up
kids_pm <- read_rds("Output/kids_pm_full.rds")

kids_pm_check <- kids_pm %>% 
  group_by(home_winter_id) %>% 
  summarise(obs = n()) %>% 
  ungroup()

ak_dt_home <- read_csv("Input/AK/ak_dt_home.csv")
nn_dt_home <- read_csv("Input/NN/nn_dt_home.csv")
wmt_dt_home <- read_csv("Input/WMT/wmt_dt_home.csv")

dt_home <- rbind(ak_dt_home, nn_dt_home, wmt_dt_home) %>% 
  rename_all(tolower) %>% 
  filter(!is.na(homewinterid)) %>% 
  mutate(home_winter_id = as.character(homewinterid)) %>% 
  left_join(kids_pm_check, by = "home_winter_id")

##### Add to new variables to Kids PM data #####
kids_pm_new <- kids_pm %>% 
  mutate(home_winter_id = as.character(home_winter_id),
         area = as.factor(area),
         pm_date = date,
         pm_time = time,
         pm_datetime = datetime,
         pm = as.numeric(pm)) %>% 
  select(home_winter_id, area, pm_date, pm_time, pm_datetime, pm) %>% 
  arrange(home_winter_id, pm_date) %>% 
  group_by(home_winter_id) %>% 
  # new variables summarizing pm2.5 within each home_winter_id
  mutate(pm_total_observations = n(),
         pm_total_summed = sum(pm),
         min_datetime = first(pm_datetime),
         max_datetime = last(pm_datetime),
         min_datetime = ymd_hms(min_datetime),
         max_datetime = ymd_hms(max_datetime),
         diff_datetime = interval(min_datetime, max_datetime),
         pm_total_days = diff_datetime/86400,
         pm_24hr_mean = pm_total_summed/pm_total_days,
         pm_rolling_24hr = rollsum(pm, 1440, fill = "extend", align = "right"),
         obs = row_number(),
         sampling_day = if_else(obs < 1441, 1, 0),
         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day),
         sampling_day = if_else(obs < 5761 & obs > 4320, 4, sampling_day),
         sampling_day = if_else(obs < 7201 & obs > 5760, 5, sampling_day),
         sampling_day = if_else(obs < 8641 & obs > 7200, 6, sampling_day),
         sampling_day = if_else(obs < 10081 & obs > 8640, 7, sampling_day),
         sampling_day = if_else(obs > 10080, 8, sampling_day)) %>%
  group_by(home_winter_id, sampling_day) %>% 
  mutate(pm_24hr_daily = sum(pm),
         pm_24hr_daily_weighted = pm_24hr_daily*(1440/n()),
         pm_sampling_day = sampling_day) %>% 
  ungroup() %>% 
  select(-min_datetime, -max_datetime, -diff_datetime, -obs, -sampling_day)

# Save pm dataset as RDS and CSV
write_rds(kids_pm_new, "Output/kids_pm_new.rds")
#write_csv(kids_pm_new, "Output/kids_pm_new.csv")
```

# Original code/files from Emily - save for future reference
```{r}
##### Load combined dusttrack files from Emily W #####
##### Compare these with the raw data pulled in above #####

### 2019-12-05 pm_summary.Rmd
PM data cleaning updates

---
title: "Kids PM Summary"
author: "Ethan Walker"
date: "Started 22 Nov 2019"
output: pdf_document
---

@@

```{r}
# Load data
kids_pm <- read_rds("Output/kids_pm.rds")
```

# Summary stats
```{r}
summary_kids_pm <- kids_pm %>% 
  summarize(mean(pm_24hr_mean), sd(pm_24hr_mean),
            min(pm_24hr_mean), max(pm_24hr_mean),
            min(pm), max(pm),
            min(pm_sampling_days), max(pm_sampling_days),
            min(pm_total_observations), max(pm_total_observations)) %>% 
  t()
summary_kids_pm
```

>>>>

---
title: "Kids PM Summary"
author: "Ethan Walker"
date: "Started 22 Nov 2019, Updated 5 Dec 2019"
output: pdf_document
---

@@

```{r}
# Load data
kids_pm <- read_rds("Output/kids_pm_new.rds")
```

# Summary stats
```{r}
summary_kids_pm <- kids_pm %>% 
  na.exclude() %>% 
  #filter(pm < 0)
  summarize(mean(pm_24hr_mean), sd(pm_24hr_mean),
            min(pm_24hr_mean), max(pm_24hr_mean),
            min(pm), max(pm), mean(pm_total_days),
            min(pm_total_days), max(pm_total_days),
            min(pm_total_observations), max(pm_total_observations),
            mean(pm_24hr_daily), mean(pm_24hr_daily_weighted)) %>% 
  t()
summary_kids_pm
```
